---
title: "tutorial_05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
**Part 1**
```{r}
library(dplyr)
olympics<-read.csv("../data/athlete_events.csv")
by_year<-olympics%>%
  filter(Season=='Summer')%>%
  arrange(Year)%>%
  group_by(Year)
q1_tbl<-summarize(by_year,n_noc=length(unique(NOC)))
q1_tbl
```

**Part 2**

Create a tibble q2_tbl that contains the number of years since the previous summer olympics. What was the longest and shortest number of years between Olympics?

Shortest number of years is 4 while longest number of years is 12
```{r}

q2_tbl<-q1_tbl%>%
  mutate(lag=lag(Year),yrs_to_past=Year-lag,.after=Year)%>%
  select(Year,yrs_to_past)
q2_tbl
#min_rank(q2_tbl$yrs_to_past)
```
**Part 3**
we need to find mean row-wise
Not quite sure why do we have to use filter from stats package

```{r}
by_year_1952<-q1_tbl%>%
  filter(Year>1950)
q3_tbl<-by_year_1952%>%
  mutate(yr_bef=lag(n_noc),yr_aft=lead(n_noc))%>%
  rowwise()%>%
  mutate(ma=mean(c_across(n_noc:yr_aft),na.rm=TRUE))
q3_tbl

```

**Part 4**

Compute the proportion of total athletes in each Games that an NOC sends.Store your output in q4_tbl.

idea: using ID to keep track number of athletes and can use .groups to change what we have grouped initially
```{r}
q4_tbl<-olympics%>%
  filter(Season=='Summer')%>%
  select(ID,Year,NOC)%>%
  group_by(Year,NOC)%>%
  summarise(noc_ath=length(unique(ID)),.groups="drop_last")%>%
  mutate(tot_ath=sum(noc_ath),prop=noc_ath/tot_ath)%>%
  ungroup()
q4_tbl
```

**Part 5**
In which year after 1950 did the USA contingent first become the majority participant of the Olympic games?

Idea:we can rank the proportions from q4 to determine the majority participant
The first year that USA contingent first become majority participant is in 1956
```{r}
q5_tbl<-q4_tbl%>%
  filter(Year>1950)%>%
  group_by(Year)%>%
  mutate(rank=dense_rank(desc(prop)))%>%
  filter(NOC=='USA',rank==1)
#To print out first occurence of USA being majority
q5_tbl[1,]
```

**Part 6**
For each athlete, compute the number of Olympics participated in before receiving his/her first medal.
If he/she never won a medal, put the value down as Inf. If he/she won a medal in their first Games,
then put the value as 0. Store your output data frame as q6_tbl.
Idea:For medals, find index of first non-NA. Look at year and rank the years. using index of first non-NA we find out the year when first medal was won.
```{r}
#function which takes in year and medal as parameters
non_na<-function(x,y){
  indexes<-which(!is.na(y))
  year_rank<-dense_rank(x)
  #meaning no medal won
  if (length(indexes)==0){
    return (Inf)
  }else{
    minimum<-min(indexes)
    return (year_rank[minimum]-1)
  }
}
q6_tbl<-olympics%>%
  filter(Season=='Summer')%>%
  group_by(ID)%>%
  arrange(Year)%>%
  summarise(num_ol_bef=non_na(Year,Medal))
q6_tbl
  
  
```
**Part 7**
Idea:First we filter out from year 2016 then we use str_extract to label which data is men,women and mixed
```{r}
library(stringr)
q7_tbl<-olympics%>%
  filter(Season=='Summer',Year==2016)%>%
  mutate(gender=str_extract(Event,"(Women)|(Men)|(Mixed)"))%>%
  filter(!gender=='Mixed')%>%
  group_by(gender,Sport)%>%
  summarise(ave_ht=mean(Height,na.rm=TRUE),ave_wt=mean(Weight,na.rm=TRUE))
q7_tbl
  
```

**Part 8**
One question that interests me is, is there any relationship between the sports played and the average height of the players(both men and female)?
More specifically, are basketball players particularly taller than other sports players


